---
title: "Configuration files for R projects"
vignette: >
  %\VignetteIndexEntry{Configuration files for R projects}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
    error: true
    message: true
---

ronfig aims to enable intuitive configuration of R projects via a single
function, `load_config()`. It's raison d'être is to handle the situation where
you have multiple, somewhat overlapping, parameter configurations that you
want to quickly switch between whilst avoiding, potentially error-inducing, copy
and paste.

To keep things simple, rather than introducing another language for our
configuration file, we restrict ourselves to a subset of base R. Currently we
allow the following set of functions/operators:

- `<-`, `=`, `+`, `-`, `*`, `:`,
- `$`, `[`, `[[`,
- `$<-`, `[<-`, `[[<-`,
- `c()`,
- `as.Date()`,
- `array(), `matrix()`,
- `list()`, `data.frame()`,
- `Sys.Date()`, `Sys.time()`.
- `seq()`, `sequence()`, and `seq_len()`.
- `file.path()`

We also define a convenience function, `cc()`, which automatically quotes input
to save typing.

This use of a reduced subset of R is primarily to ensure that loading a
configuration file has no side-effects on a users environment and performs
limited computation. Under the hood, `load_config()` is little more than careful
wrapping of `base::sys.source()` and `utils::modifyList()` with care taken as to
where evaluation takes place.

## Example

### Configuration file

Configuration files look similar to the following example file that we bundle in
the package:

```{r, comment=''}
file <- system.file("config.R", package = "ronfig")
cat(readChar(file, file.info(file)$size))
```

### Loading

We can easily load any of the example configurations in to a list in our local
environment:

```{r}
library(ronfig)
str(load_config(file))
str(load_config(file, "debug"))
str(load_config(file, "forecast"))
```

### Using other functions

To make other functions available within the configuration file, users can
inject `carrier::crate()` objects in to the evaluation environment. As an
example, assume we wanted to permit the use of `mean()` within our configuration
file. By default this would error:

```{r}
f <- tempfile()
cat("default <- list(a=mean(1:10))", file = f)
with(load_config(f), a)
```
If we now inject `mean()` as a crated function, things work as desired:

```{r}
crate <- carrier::crate(function(x) mean(x))
with(load_config(f, crates = list(mean = crate)), a)
unlink(f)
```

## Related work

- [config](https://cran.r-project.org/package=config) which uses yaml for
  it's configuration file.

