---
title: "Configuration files for R projects"
output:
    html:
        meta:
            css: ["@default@1.13.67", "@callout@1.13.67", "@article@1.13.67"]
            js: ["@sidenotes@1.13.67", "@copy-button@1.13.67", "@callout@1.13.67", "@toc-highlight@1.13.67"]
        options:
            toc: true
            js_highlight:
                package: prism
                version: 1.29.0
        
vignette: >
  %\VignetteEngine{litedown::vignette}
  %\VignetteIndexEntry{ronfig}
---

```{r, include = FALSE}
litedown::reactor(print = NA)
```

ronfig aims to enable intuitive configuration of R projects via a single
function, `load_config()`. It's raison d'être is to handle the situation where
you have multiple, somewhat overlapping, parameter configurations that you
want to quickly switch between whilst avoiding, potentially error-inducing, copy
and paste.

To keep things simple, rather than introducing another language for our
configuration file, we restrict ourselves to a subset of base R. Currently we
allow the following set of functions/operators:

- <-, =, +, -, *, :,
- $, [, [[,
- $<-, [<-, [[<-,
- c,
- as.Date,
- array, matrix,
- list, data.frame,
- Sys.Date, Sys.time.
- seq, sequence, and seq_len.

We also define a convenience function, `cc`, which automatically quotes input
to save typing.

::: {.callout-note data-legend="Implementation"}
This use of a reduced subset of R is primarily to ensure that loading a
configuration file has no side-effects on a users environment and performs
limited computation. Under the hood, `load_config()` is little more than careful
wrapping of `base::sys.source()` and `utils::modifyList()` with care taken as to
where evaluation takes place.
:::

## Example

### Configuration file

Configuration files look similar to the following example file that we bundle in
the package:

```{r, comment=''}
file <- system.file("config.R", package = "ronfig")
cat(readChar(file, file.info(file)$size))
```

### Loading
We can easily load any of the example configurations in to a list in our local
environment:

```{r}
library(ronfig)
str(load_config(file))
str(load_config(file, "debug"))
str(load_config(file, "forecast"))
```

## Related work

- [config](https://cran.r-project.org/package=config) which uses yaml for
  it's configuration file.
